### 2014-02-28 paper/retrofit.tex

While we would like to add information flow control to \emph{any}
language, we will have to make some assumptions about the form of the
target language.  We consider languages whose semantics can be expressed
in the form of abstract machines |(tS, tE, te, tv, ->)|, where |tS| is the
global machine state, |tE| is an evaluation context, |te| is an
expression in the language, |tv| is the set of values, and |->| is the reduction relation on such
configuration, i.e., $|->| : (|tS| \times |tE| \times |te|)
\rightharpoonup (|tS| \times |tE| \times |te|)$.  Metavariables in this
target language have been colored red.

In Figure~\ref{fig:ml}, we give an example of a standard ML-like
language with mutable references.  These semantics have been prepared
for interoperability in the style of Matthews and Findler~\Red{XXX}; in
particular, the evaluation rules have been expressed in terms of an
evaluation context $\mathcal{E}_{\tar{\Sigma}}$.  Without the embedding,
this evaluation context is defined as |conf tS te @= tS, tE[ te ]|, but
when we perform the embedding, we will reinterpret the meaning
of this evaluation context, lifting these semantics to an IFC setting.

@@

Next, we describe the IFC language.  \Red{Description of the IFC language as
a standard floating label system, or backwards reference to the
background section.}  In Figure~\ref{fig:ifc}, we describe the syntax
of the language, and provide rules for all operations which operate on a single
thread.  \Red{One could imagine these rules giving the definition of a very
boring information flow control language, by defining |niconf iS id il ie| appropriately
and providing an appropriate erasure function.}

We use the notation |filter il Q| to indicate a filter function which removes message whose labels do not
flow to |il|. Specifically, if |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]

>>>>

Our overall plan will be to take a \textbf{{\color{red} target language}} with a formal
operational semantics and combine it with an \textit{{\color{blue} information
flow control language}}.  For concreteness, we use a simple, untyped,
ML-like language with mutable references as our target language.
However, our technique will be indifferent to the actual target language
chosen: it need only be prepared for interoperability by being expressed
in the style of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220}.  This procedure will
equip the target language with primitives for implementing dynamic
information control in the style of LIO~\cite{lio}, which we will
describe in more detail shortly.

We have typeset nonterminals of the target language using \textbf{{\color{red}
bold font}} while the nonterminals of the IFC language have been typeset
with \textit{{\color{blue} italic font}}.  Readers are encouraged to view
a color copy of this paper, where target language nonterminals are colored red
while IFC language nonterminals are colored blue.

\subsection{Target language: Mini-ML}

In Figure~\ref{fig:ml}, we give a simple, untyped ML-like language with
mutable references \Red{and errors}, prepared for combination with an
information flow control language.  The presentation is standard, but it
is worth paying some attention to the Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} used to define the
operational semantics of the system.  Our language defines an evaluation
context |tE|; however, the evaluation rules have been expressed in terms
of a different evaluation context $\mathcal{E}_{\tar{\Sigma}}$.  To
derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, tE[ te ]|.
However, when we combine this language with an IFC language, we must
reinterpret the meaning of this evaluation context.
In general, we require that a target language be expressed in terms
of some global machine state |tS|, some evaluation context |tE|,
some expressions |te|, some set of values |tv| and a reduction
relations on full configurations $|tS| \times |tE| \times |te|$.

@@

\subsection{IFC language}

Information flow control seeks to enforce a security policy known as
\emph{noninterference}: public results must not depend on secret inputs.
Modern, dynamic IFC systems usually achieve this by associating a \emph{label}
with data, which encodes a policy of whether or not data tagged with one
label can flow to |canflowto| another label.  Since our embedding is agnostic
to the choice of labeling scheme, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

In Figure~\ref{fig:ifc}, we give the syntax and \emph{single-threaded}
evaluation rules for a minimal information-flow control language.  The
evaluation context $\mathcal{E}_{\ifc{\Sigma}}$ represents the context
of the current thread which is running, including on the right-hand side
the current label $l$ (which protects the program counter of the thread)
and the thread ID $i$.

Ordinarily, information flow control languages are defined by directly
stating a base language plus information flow control operators.  In
contrast, our language is purposely minimal: it does not have sequencing
operations, control flow, or other constructs.  However, it contains
support for the following core information flow control features:

\begin{itemize}
    \item First-class labels, with label values $l$ as well as operations for computing on
labels (|canflowto|, |lub| and |glb|),
    \item Operations for inspecting (\textbf{getLabel}) and modifying (\textbf{setLabel}) the current label of the thread (note a thread can only increase its label, and not decrease it),
    \item Operations for non-blocking inter-thread communication (\textbf{send} and \textbf{recv}),
        which interact with the global store of per-thread message queues |iS|, and
    \item A forking operation, which will be described in more detail in the next section.
\end{itemize}

The form of \textbf{send}/\textbf{recv} are inspired by methods of inter-worker communication
in web browser APIs \Red{cite}, however in Section~\ref{sec:extensions} we
discuss other possible choices for information flow control aware communication
primitives.  Their rules are non-standard and worth commenting on: when a thread places
a message on the message queue of the thread it is sending to, it also adds the current
label when it sent the message.  When the recipient reads out messages from its queue,
it filters out (notated using |filter il Q|) messages whose labels do not
flow to |il|. Specifically, if |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|; otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
